---
title: "CiC Rejection Sampling"
output: html_notebook
---

This workbook is responsible for creating some useful model inputs:
* Creating a target duration distribution for each join age using survival analysis
* Accounting for the effect of aging out (a dynamic which must be handled separately from survival analysis)
* Creating model input files containing pre-calculated CiC periods which when sampled match the inferred target duration distribution


```{r}
library(lubridate)
library(survival)
library(survminer)
library(stringr)
library(reshape2)
library(zoo)
library(dplyr)
library(tidyr)
```

```{r}
source("helpers.R")
```

```{r}

extract_date <- as.Date("2020-03-31")
input_csv <- "../data/scc/2021-04-08/suffolk-scrubbed-episodes-20210219.csv"
# All of the periods we have simulated
periods_universe_csv <- "/Users/Seb/code/witan.csc.suffolk/data/outputs/generated-candidates.csv"

output_csv <- "../data/scc/2021-04-08/target-distribution.csv"
age_out_proportions_output_csv <- "../data/scc/2021-04-08/age-out-proportions.csv"

simulated_output_csv <- "../data/scc/2021-04-08/simulated-candidates.csv"
projected_output_csv <- "../data/scc/2021-04-08/projected-candidates.csv"
simulated_age_out_output_csv <- "../data/scc/2021-04-08/simulated-age-out-candidates.csv"
projected_age_out_output_csv <- "../data/scc/2021-04-08/projected-age-out-candidates.csv"

data <- read.csv(input_csv)
data$report_date <- ymd(data$report_date)
data$ceased <- ymd(data$ceased)

data.frame(xs = c(data$report_date, data$ceased)) %>%
  filter(!is.na(xs)) %>%
  filter(xs > as.Date("2020-01-01")) %>%
  ggplot(aes(xs)) + geom_histogram()


# Fix dodgy NAs

# data$ceased[data$ceased > extract_date] <- NA
# data <- data[data$report_date <= extract_date,]
#
# data %>% filter(period_id == "881-1")
# data %>% filter(period_id == "486-1")
#
# data %>% filter(period_id == "1815-1")
```

```{r}

# Creates data that will be fed to the survival curve fitting model.
# We only know which month each child's birthday falls in, so we create a record
# for each day of this month.

noise <- 7

survival_data <- data %>%
  mutate(report_date = ymd(report_date),
         ceased = ymd(ceased)) %>%
  group_by(period_id) %>%
  dplyr::summarise(period_start = min(report_date),
                   period_end = max(ceased),
                   birth_month_beginning = ymd(paste0(DOB[1], "-01")),
                   birth_month_end = birth_month_beginning + months(1) - days(1))

# We think the child aged out if they *could* have hit 18 years - 1 day given their birthday range
# That means if we assume the oldest they could be, did that cross 18 years - 1 day.
# That means if we take their earlier birthday, is it *at least* 18 years - 1 day.

survival_data2 <- survival_data %>%
  mutate(birthday_lower = pmax(birth_month_beginning, eighteen_years_before(replace_na(period_end, extract_date))),
         birthday_upper = pmin(birth_month_end, period_start),
         # aged_out = year_diff(birth_month_beginning, replace_na(period_end, extract_date)) >= 18,
         # Removing this period_end assignment - don't think we need it
         # period_end = if_else(aged_out, eighteen_years_after(birthday_upper) - days(1), period_end),
         event = !is.na(period_end),
         duration = if_else(event,
                            day_diff(period_start, period_end),
                            day_diff(period_start, extract_date)))

# These cases violate one of more of our assumptions. We'll assume they left at 18.
survival_data2 %>% filter(birthday_lower > birthday_upper)

survival_data3 <- survival_data2 %>%
  mutate(birthday_lower = pmin(birthday_lower, birthday_upper) # Only matters if there are age 18+ leavers above
         ) %>%
  inner_join(data.frame(n = 1:100), by = character()) %>%
  mutate(birthday = as.Date(impute_birthday(birthday_lower, birthday_upper), origin = "1970-01-01"),
         admission_age = as.factor(clamp(year_diff(birthday, period_start), 0, 17)),
         max_duration = day_diff(period_start, eighteen_years_after(birthday)),
         duration = pmin(duration, max_duration),
         fuzzed_duration = binomial_noise(duration, max_duration, noise)
         ) %>%
  as.data.frame

survival_data4 <- survival_data3 %>%
  mutate(aged_out = year_diff(birthday, period_start + days(fuzzed_duration)) >= 17, # Going to say no one leaves aged 17
         event = event | aged_out, # All aged out children are leavers
         duration = if_else(aged_out, max_duration, duration),
         fuzzed_duration = if_else(aged_out, max_duration, fuzzed_duration),
         fuzzed_exit_age = fuzzed_duration + day_diff(birthday, period_start))

```



```{r}
fit <- survfit(Surv(fuzzed_duration, event) ~ admission_age, data = survival_data4 %>% filter(!aged_out))
quantiles <- reshape2::melt(stats::quantile(fit, probs = seq(0,0.9999,length.out = 10000))$quantile) %>%
  mutate(join_age = as.integer(str_replace(Var1, "admission_age=", ""))) %>%
  select(-1) %>%
  as.data.frame
colnames(quantiles) <- c("quantile", "duration", "admission_age")

quantiles <- dcast(admission_age ~ quantile, data = quantiles, drop = FALSE, value.var = "duration")
for (row in 1:nrow(quantiles)) {
  # I expect there's a better way of doing this
  age <- row - 1
  max_days <- (18 - age) * 365
  quantiles[row,is.na(quantiles[row,])] <- max_days
  quantiles[row,quantiles[row,] > max_days] <- max_days
}
quantiles <- cbind(quantiles, data.frame(`100` = 17:1*365))
colnames(quantiles) <- c("admission_age", seq(0, 10000, length.out = ncol(quantiles) - 1))
quantiles <- melt(quantiles, id.vars = "admission_age") %>%
  mutate(quantile = as.numeric(variable) / 10000.0,
         duration = value) %>%
  select(admission_age, duration, quantile) %>%
  group_by(admission_age, duration) %>%
  summarise(quantile = max(quantile))

target_distribution <- quantiles %>%
  mutate(duration_group = duration %/% 7) %>%
  group_by(admission_age, duration_group) %>%
  summarise(quantile = max(quantile)) %>%
  arrange(admission_age, duration_group) %>%
  mutate(density = quantile - coalesce(lag(quantile), 0)) %>%
  mutate(density = density * 100 / sum(density))



write.csv(target_distribution, file = output_csv, row.names = FALSE)
```

```{r}

# Calculate probability that joiner of each age makes it past 17th birthday

fit <- survfit(Surv(fuzzed_exit_age, event) ~ admission_age, data = survival_data4)
quantiles <- reshape2::melt(stats::quantile(fit, probs = seq(0,0.9999,length.out = 10000))$quantile) %>%
  mutate(join_age = as.integer(str_replace(Var1, "admission_age=", ""))) %>%
  select(-1) %>%
  as.data.frame
colnames(quantiles) <- c("quantile", "exit_age", "admission_age")

age_out_proportions <- quantiles %>%
  group_by(admission_age) %>%
  mutate(age_out_quantile = min(if_else(exit_age > 17 * 365.25, quantile, NA_real_), na.rm = TRUE)) %>%
  mutate(p = age_out_quantile - quantile) %>%
  group_by(admission_age, exit_age) %>%
  summarise(p = (100 - min(p)) / 100.0) %>%
  rename(current_age_days = exit_age)

write.csv(age_out_proportions, file = age_out_proportions_output_csv, row.names = FALSE)

```

```{r}
library(RSQLite)
db <- dbConnect(RSQLite::SQLite(), "mastodon.sqlite")
```

```{r}
# Upload periods universe
periods_universe <- read.csv(periods_universe_csv)

staged <- periods_universe
colnames(staged) <- c("provenance", "id", "sample_index", "admission_age", "admission_age_days", "duration", "episodes_edn", "aged_out")
dbWriteTable(db, "periods", staged, overwrite = TRUE, row.names = FALSE)

# Upload target distribution
staged <- target_distribution[,c("admission_age", "duration_group", "density")]
dbWriteTable(db, "target_distribution", staged, overwrite = TRUE, row.names = FALSE)
```

We create versions of the input periods to try and get good coverage of the whole target distribution.
This means adjusting the total duration by up to 10 weeks either way.
We don't filter periods which go over 18 because we adjust the birthdays in the following section.

```{r}
noise <- data.frame(x = (seq(-10, 10) * 7))
dbWriteTable(db, "noise", noise, overwrite = TRUE, row.names = FALSE)
```

```{sql connection=db}
DROP TABLE IF EXISTS fuzzy_periods;
```

```{sql connection=db}
CREATE TABLE fuzzy_periods AS
WITH x AS (
  SELECT x FROM noise
)
SELECT p.provenance, p.id, p.sample_index, p.admission_age, p.admission_age_days,
  cast(p.duration + x AS int) AS duration, CAST(duration + x AS int) / 7 AS duration_group, episodes_edn
FROM periods AS p
  CROSS JOIN x
WHERE (duration + x >= 28 OR x = 0)
AND aged_out = 'false';
```

```{sql connection=db}

INSERT INTO fuzzy_periods
WITH x AS (
  SELECT x FROM noise
)
SELECT p.provenance, p.id, p.sample_index, p.admission_age, p.admission_age_days,
  cast(p.duration + x AS int) AS duration, CAST(duration + x AS int) / 7 AS duration_group, episodes_edn
FROM periods AS p
  CROSS JOIN x
WHERE duration + x >= 0
AND duration + x < 28
AND x != 0
AND aged_out = 'false';
```

For all those who join aged at least 1, we create versions of all simulated and projected data with fuzzed birthdays: up to 5 weeks in either direction.
```{r}
noise <- data.frame(x = (seq(-5, 5) * 7))
dbWriteTable(db, "noise", noise, overwrite = TRUE, row.names = FALSE)
```

```{sql connection=db}
WITH x AS (
  SELECT x FROM noise
)
INSERT INTO fuzzy_periods
SELECT p.provenance, p.id, p.sample_index, CAST(p.admission_age_days + x AS int) / 365, CAST(p.admission_age_days + x AS int) AS admission_age_days, p.duration, p.duration_group, episodes_edn
FROM fuzzy_periods AS p
CROSS JOIN x
WHERE x != 0 AND CAST(p.admission_age_days + x AS int) / 365 >= 1
AND provenance IN ('S', 'P');
```
```{sql connection=db}
DROP TABLE IF EXISTS candidate_distributions;
```

```{sql connection=db}
DROP TABLE IF EXISTS target_projected_distribution;
```

```{sql connection=db}
CREATE TABLE candidate_distributions AS
SELECT provenance, admission_age, duration_group, COUNT(*) AS density
FROM fuzzy_periods
GROUP BY provenance, admission_age, duration_group;
```


```{sql connection=db}
CREATE TABLE target_projected_distribution AS
WITH open_period_durations AS (
  SELECT *, CAST(duration AS int) / 7 AS duration_group
  FROM periods
  WHERE provenance = 'C'
),
tpd AS (
  SELECT pp.admission_age, td.duration_group, CASE WHEN pp.duration_group = td.duration_group THEN density * (7 - (pp.duration % 7)) ELSE density END AS density
  FROM open_period_durations pp
  INNER JOIN target_distribution td
  ON pp.admission_age = td.admission_age
  AND td.duration_group >= pp.duration_group
)
SELECT admission_age, duration_group, SUM(density) AS density
FROM tpd
GROUP BY admission_age, duration_group;
```

We need to ensure that every ID to be projected has at least one period with a non-zero density in the target (otherwise it will never be chosen)

```{r}

target_projected_duration_groups <- dbGetQuery(db, "SELECT admission_age, duration_group, density
                                  FROM target_projected_distribution")

for (age in 0:16) {
  print(ggplot(target_projected_duration_groups %>% filter(admission_age == age), aes(duration_group, density)) +
    geom_bar(stat = "identity") +
    facet_wrap(vars(admission_age), scales = "free_y"))
}

non_join_duration_groups <- dbGetQuery(db, "WITH empty_ids AS (
SELECT fp.id
FROM fuzzy_periods fp
LEFT OUTER JOIN target_projected_distribution pd
ON fp.admission_age = pd.admission_age
AND fp.duration_group = pd.duration_group
WHERE fp.provenance = 'P'
GROUP BY fp.id
HAVING SUM(coalesce(pd.density, 0)) = 0
)SELECT id, admission_age, duration_group, COUNT(*) AS density
FROM fuzzy_periods
WHERE id IN (SELECT id FROM empty_ids)
AND provenance = 'P'
GROUP BY id, admission_age, duration_group;")

combined_duration_groups <- rbind(cbind(id = "target", target_projected_duration_groups),
      non_join_duration_groups) %>%
  group_by(id, admission_age) %>%
  mutate(density = density / max(density))

library(ggthemes)

for (age in 0:17) {
  print(ggplot(combined_duration_groups %>% filter(admission_age == age), aes(duration_group, density, fill = id)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(vars(admission_age)) +
  scale_fill_manual(values = tableau_color_pal("Tableau 20")(20)))

}

## Shows they should all remain until 18

```


```{sql connection=db}
DROP TABLE IF EXISTS simulated_candidates;
```

```{sql connection=db}
DROP TABLE IF EXISTS projected_candidates;
```

```{sql connection=db}
-- Update destination if necessary
CREATE TABLE projected_candidates AS
WITH source_periods AS (
  -- Update source periods if necessary
  SELECT *
  FROM fuzzy_periods
  WHERE provenance = 'P'
),
target_distribution AS (
  -- Update target if necessary
  SELECT * FROM target_projected_distribution
),
candidate_distribution AS (
  SELECT admission_age, duration_group, COUNT(*) AS density
  FROM source_periods
  GROUP BY admission_age, duration_group
),
reject_proportions AS (
  SELECT t.admission_age, t.duration_group, t.density / c.density AS reject_ratio
  FROM target_distribution t
  LEFT JOIN candidate_distribution c
  ON c.admission_age = t.admission_age
  AND c.duration_group = t.duration_group
)
SELECT fp.*, rp.reject_ratio
FROM source_periods fp
INNER JOIN reject_proportions rp
ON fp.admission_age = rp.admission_age
AND fp.duration_group = rp.duration_group
WHERE random() <= 0.25;
```

```{sql connection=db}
-- If no samples from a particular candidate made it through rejection sampling,
-- bypass rejection sampling to add them in, because otherwise they will be dropped completely.
INSERT INTO projected_candidates
WITH missing_periods AS (
SELECT *, row_number() OVER (PARTITION BY id ORDER BY random()) AS rand_id
FROM fuzzy_periods fp
WHERE fp.provenance = 'P'
AND id NOT IN (
  SELECT DISTINCT id
  FROM projected_candidates
)
)
-- Maybe we ought to set the duration to cause them to age out at 18 - this seem like their destiny.
-- However this would cause them to remain in their final placement for a long time.
SELECT provenance, id, sample_index, admission_age, admission_age_days, duration, duration_group, episodes_edn, 1.0 AS reject_ratio
FROM missing_periods
WHERE rand_id <= 100;
```

```{sql connection=db}
-- Update destination if necessary
CREATE TABLE simulated_candidates AS
WITH source_periods AS (
  -- Update source periods if necessary
  SELECT *
  FROM fuzzy_periods
  WHERE provenance = 'S'
),
tgd AS (
  -- Update target if necessary
  SELECT * FROM target_distribution
),
candidate_distribution AS (
  SELECT admission_age, duration_group, COUNT(*) AS density
  FROM source_periods
  GROUP BY admission_age, duration_group
),
reject_proportions AS (
  SELECT t.admission_age, t.duration_group, t.density / c.density AS reject_ratio
  FROM tgd t
  LEFT JOIN candidate_distribution c
  ON c.admission_age = t.admission_age
  AND c.duration_group = t.duration_group
)
SELECT fp.*, rp.reject_ratio
FROM source_periods fp
INNER JOIN reject_proportions rp
ON fp.admission_age = rp.admission_age
AND fp.duration_group = rp.duration_group
WHERE random() <= 0.25;

```


```{r}

staged <- dbGetQuery(db, "SELECT * FROM projected_candidates");
write.csv(staged, file = projected_output_csv, row.names = FALSE)

staged <- dbGetQuery(db, "SELECT * FROM simulated_candidates");
write.csv(staged, file = simulated_output_csv, row.names = FALSE)

staged <- dbGetQuery(db, "SELECT provenance, id, sample_index, admission_age, admission_age_days, duration, episodes_edn
FROM periods
WHERE provenance = 'S'
AND aged_out = TRUE
AND admission_age_days + duration > (17 * 365)")
write.csv(staged, file = simulated_age_out_output_csv, row.names = FALSE)


staged <- dbGetQuery(db, "SELECT provenance, id, sample_index, admission_age, admission_age_days, duration, episodes_edn
FROM periods
WHERE provenance = 'P'
AND aged_out = TRUE
AND admission_age_days + duration > (17 * 365)")
write.csv(staged, file = projected_age_out_output_csv, row.names = FALSE)

```
